cardinalR: Generating interesting high-dimensional data structures

### 2025-06-24 paper-cardinalR.Rmd
Added Mobius structure

The `gen_longlinear(n, p)` function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise. Each variable $X_i$ is created as $X_i = \text{scale}_i \cdot \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i$, where $\text{scale}_i \sim U(-10, 10)$ is randomly chosen direction and stretch, $\text{shift}_i \sim U(-300, 300)$ is large offset to spread features apart, and $\epsilon \sim N(0, 0.03 \cdot n)$ is Gaussian noise.

@@

The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p$-dimensional space. This classical non-orientable surface loops back on itself with a half-twist.

The Mobius strip is generated using `geozoo::mobius(n, p = 3)$points`, which samples $n$ points from a parametrization of the surface defined by angle $\theta$ and strip width $w$. The cartesian coordinates are;

$$
X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta),
$$
$$
X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta),
$$
$$
X_3 = \frac{w}{2} \sin(\theta/2).
$$

This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable 1-sided surface.

For $p > 3$, the function appends $p - 3$ dimensions with small Gaussian noise:

$$
x_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
$$

>>>>

The `gen_longlinear(n, p)` function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise. Each variable $X_i$ is created as $X_i = \text{scale}_i \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i$, where $\text{scale}_i \sim U(-10, 10)$ is randomly chosen direction and stretch, $\text{shift}_i \sim U(-300, 300)$ is large offset to spread features apart, and $\epsilon \sim N(0, 0.03n)$ is Gaussian noise.

@@

The `gen_mobius(n, p)` function generates a dataset of $n$ points that form a Mobius strip embedded in the first three dimensions of a $p\text{-}D$. This classical non-orientable surface loops back on itself with a half-twist.

The Mobius strip is generated using `geozoo::mobius(n, p = 3)$points`, which samples $n$ points from a parametrization of the surface defined by angle $\theta$ and strip width $w$. The cartesian coordinates are $X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta)$, $X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta)$, and $X_3 = \frac{w}{2} \sin(\theta/2)$. This maps a $2\text{-}D$ band with a half-twist into $3\text{-}D$ space, forming a non-orientable one-sided surface. For $p > 3$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $3\text{-}D$ Mobius into $p\text{-}D$, where $j = 4, \dots, p$.

### 2025-06-24 paper-cardinalR.tex
Added Mobius structure

The \texttt{gen\_longlinear(n,\ p)} function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise. Each variable \(X_i\) is created as \(X_i = \text{scale}_i \cdot \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i\), where \(\text{scale}_i \sim U(-10, 10)\) is randomly chosen direction and stretch, \(\text{shift}_i \sim U(-300, 300)\) is large offset to spread features apart, and \(\epsilon \sim N(0, 0.03 \cdot n)\) is Gaussian noise.

@@

The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\)-dimensional space. This classical non-orientable surface loops back on itself with a half-twist.

The Mobius strip is generated using \texttt{geozoo::mobius(n,\ p\ =\ 3)\$points}, which samples \(n\) points from a parametrization of the surface defined by angle \(\theta\) and strip width \(w\). The cartesian coordinates are;

\[
X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta),
\]
\[
X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta),
\]
\[
X_3 = \frac{w}{2} \sin(\theta/2).
\]

This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable 1-sided surface.

For \(p > 3\), the function appends \(p - 3\) dimensions with small Gaussian noise:

\[
x_j \sim N(0, 0.05^2), \quad j = 4, \dots, p.
\]

>>>>

The \texttt{gen\_longlinear(n,\ p)} function simulates a high-dimensional linear structure with a dominant linear trend and small additive noise. Each variable \(X_i\) is created as \(X_i = \text{scale}_i \left(0, 1, \dots, n{-}1 + \epsilon\right) + \text{shift}_i\), where \(\text{scale}_i \sim U(-10, 10)\) is randomly chosen direction and stretch, \(\text{shift}_i \sim U(-300, 300)\) is large offset to spread features apart, and \(\epsilon \sim N(0, 0.03n)\) is Gaussian noise.

@@

The \texttt{gen\_mobius(n,\ p)} function generates a dataset of \(n\) points that form a Mobius strip embedded in the first three dimensions of a \(p\text{-}D\). This classical non-orientable surface loops back on itself with a half-twist.

The Mobius strip is generated using \texttt{geozoo::mobius(n,\ p\ =\ 3)\$points}, which samples \(n\) points from a parametrization of the surface defined by angle \(\theta\) and strip width \(w\). The cartesian coordinates are \(X_1 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \cos(\theta)\), \(X_2 = \left(1 + \frac{w}{2} \cos(\theta/2)\right) \sin(\theta)\), and \(X_3 = \frac{w}{2} \sin(\theta/2)\). This maps a \(2\text{-}D\) band with a half-twist into \(3\text{-}D\) space, forming a non-orientable one-sided surface. For \(p > 3\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(3\text{-}D\) Mobius into \(p\text{-}D\), where \(j = 4, \dots, p\).
